Actuator line modeling of vertical-axis turbines

### 2016-05-13 paper.tex
Remove more programming details

NREL has developed and released an actuator line modeling library,
SOWFA~\cite{Churchfield2014b}, for simulating horizontal-axis wind turbine
arrays using the OpenFOAM finite volume CFD library. OpenFOAM is free,
open-source, widely used throughout industry and academia, and has grown a very
active support community around itself. Though SOWFA is also open-source, and
has the ability to be coupled to NREL's FAST HAWT BEM code, its mostly
procedural style would have required significant effort and duplicate code to
adapt for cross-flow turbines. Thus, a new and more general ALM library was
developed from the ground up that could model both cross- and axial-flow
turbines, as well as standalone actuator lines. The actuator line model
developed here, dubbed turbinesFoam \cite{Bachant2016-turbinesFoam-v0.0.7}, was
also written as an extension library for OpenFOAM, and was developed freely and
openly from its inception in order to increase community engagement and research
efficiency.

turbinesFoam was written in OpenFOAM's style, using OpenFOAM's
\texttt{fvOptions} framework for adding source terms to equations at runtime.
This implementation allows the CFT-ALM to be added to many of the standard
solvers included in OpenFOAM without modification, i.e., it can be readily used
with RANS or LES, multiphase models (e.g., for simulating the free surface in
MHK installations), and even with heat transfer.

OpenFOAM and turbinesFoam are written in the C++ programming language, which
follows the object oriented programming paradigm. This characteristic helped
modularize the ALM code for increased readability and reuse. In turbinesFoam, a
turbine is a software object that is composed of actuator line objects, which
themselves are composed of actuator line element objects. Structuring the code
this way allows isolation and reuse of the functionality of individual
components. For example, the actuator line object was written such that is could
be used outside the turbine context to ensure it produces the correct forcing,
without adding the complexity of rotation, other actuator lines, etc. that would
be present in a turbine rotor. The very same actuator line objects can be used
in both axial-flow and cross-flow rotors, without having to copy code from one
to the other.

>>>>

NREL has developed and released an actuator line modeling library,
SOWFA~\cite{Churchfield2014b}, for simulating horizontal-axis wind turbine
arrays using the OpenFOAM finite volume CFD library. OpenFOAM is free,
open-source, widely used throughout industry and academia, and has grown a very
active support community around itself. Though SOWFA is also open-source, and
has the ability to be coupled to NREL's FAST HAWT BEM code, its mostly
procedural style would have required significant effort and duplicate code to
adapt for cross-flow turbines. Thus, a new and more general ALM library was
developed from the ground up that could model both cross- and axial-flow
turbines, as well as standalone actuator lines. The actuator line model
developed here, dubbed turbinesFoam \cite{Bachant2016-turbinesFoam-v0.0.7}, was
also written as an extension library for OpenFOAM, using its \texttt{fvOptions}
framework for adding source terms to equations at runtime. This implementation
allows the CFT-ALM to be added to many of the standard solvers included in
OpenFOAM without modification, i.e., it can be readily used with RANS or LES,
multiphase models (e.g., for simulating the free surface in MHK installations),
and even with heat transfer.
